100 Years of Social Work Research: A Data Science Perspective

### 2015-01-14 Analysis.Rmd
Updated files. All functional. Growth of science compared to social work functional.

Data collection date:  January 8, 2014

Database: PsychInfo (via EbscoHost Platform)

Search terms:  SO "social work" or SO "social welfare" or SO "social casework" or "social case work" or SO "social service" or "social services" or SO "human service" or SO "human services" or SO "social development" or SO "social environment"

Search limiters:  Publication Type: Peer Reviewed Journal; Document Type: Journal Article

Search modes: Boolean/Phrase

Search results were exported from EbscoHost Platformon in a _generic bibliographic format_.  This is an unstructured text file (*.txt) that was processed using in R.

```{r initialization_dataprocessing, warning=FALSE, message=FALSE, eval=FALSE}
# Clear workspace
rm(list=ls())

@@

```{r initialization_analysis, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
rm(list=ls())
setwd("/Users/beperron/Git/SocialWorkResearch/Data")
source("/Users/beperron/Git/BibWrangleR/functions/ggsurv.R")
load("piArticles.R")
library(dplyr)
library(ggplot2)
library(gridExtra)
library(survival)
library(grid)
library(png)

@@

Note the exclusion of social work today.

```{r echo=FALSE}
pi.df <- pi.df %>% filter(
        record != "Journal of Applied Social Sciences" &
        record != "Early Child Development and Care" &
        record != "The Clinical Supervisor" &
        record != "Children and Youth Services Review" &
        record != "General Hospital Psychiatry" &
        record != "Canadian Journal on Aging" &
        record != "Canadian Journal of Community Mental Health" &
        record != "Behavior Modification" &
        record != "Employee Assistance Quarterly" &
        record != "Journal of Applied Behavioral Science" & 
        record != "The Scientific Review of Mental Health Practice: Objective Investigations of Controversial and Unorthodox Claims in Clinical Psychology, Psychiatry, and Social Work" &
        record != "Systems Research and Behavioral Science" & 
        record != "Journal of Applied Rehabilitation Counseling" & 
        record != "Journal of Educational & Psychological Consultation" &
        record != "Journal of Community & Applied Social Psychology" &                                          
        record != "Evaluation and Program Planning" &                                                                  
        record != "Journal of Nonverbal Behavior" &
        record != "Brain and Cognition" &
        record != "Early Education and Development" & 
        record != "American Journal of Preventive Medicine" &
        record != "Social Work Today")
```

This section provides further cleaning of retained journals, namely merging two journals with abbreviated titles:

@@

pi.df$record <- ifelse(pi.df$record == "Practice", "Practice: Social Work in Action", pi.df$record)
pi.df$articleID <- ifelse(pi.df$articleID == 74, 42, pi.df$articleID)

pi.df$record <- ifelse(pi.df$record == "Eugenics & Social Welfare Bull.", "Eugenics and Social Welfare Bulletin", pi.df$record)
pi.df$articleID <- ifelse(pi.df$articleID == 28, 32, pi.df$articleID)

@@

number.journals
journals.unique
```

```{r unique_titles_year, message=FALSE, comment=NA, warning=FALSE, fig.height=2, fig.height=4, echo=FALSE}
journals.year <- tbl_df(pi.df)

@@

journal.count <- ggplot(n.journals.year, aes(as.numeric(year), y=n, group=1)) + 
    geom_line(colour="black") +
    #geom_point(colour="red") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    xlab("Publication year") + 
    ylab("Number") + 
    ggtitle("Number of Journals Over Time") + 
    scale_x_continuous(breaks=seq(1914, 2014, 5))

@@

@@

@@

@@

@@

@@

+ social justice
+ evidence-based practice" | "evidence based practice" | "EBP"
+ bullying
+ child welfare

@@

select.keywords <- c("depression", "hiv")
select.keywords.df <- kp.year %>% filter(tolower(keyword) %in% select.keywords)

@@

# Location of Studies

@@

#img2 <- readPNG("/Users/beperron/Git/SocialWorkResearch/Chloro.png")
#grid.raster(img2)
```

@@

@@

ggplot(location.year.top, aes(x = year, y = log_n, colour = country )) + 
    geom_line(size = 1.25) +
    ggtitle("Log Number of Articles by Countries With Highest Scholarly Ouput") +
    xlab("Publication Year") + 
    ylab("log N+1") +
    theme(legend.position=c(.210, .7))

@@

ggplot(location.year.top.no.us, aes(x = year, y = cumsum_n, colour = country )) + geom_line(size=1.00) +
    ggtitle("Cumulative Number of Article by Countries with Highest Scholarly Output \n(Excluding the United States)") +
    theme(legend.position=c(.210, .6))
```

@@

This section is currently not functional
```{r countries_not_represented, echo=FALSE}
#Tableau.countries <- read.csv("/Users/beperron/Git/SocialWorkResearch/SupportingDocs/Tableau\ Countries.csv")
```

>>>>

# Data collection

+ Article record retrieval date:  January 8, 2015
Database: PsychInfo and Social Science Index (via EbscoHost Platform)

+ Search terms:  SO "social work" or SO "social welfare" or SO "social casework" or "social case work" or SO "social service" or "social services" or SO "human service" or SO "human services" or SO "social development" or SO "social environment"

+ Search limiters:  Publication Type: Peer Reviewed Journal; Document Type: Journal Article

+ Search modes: Boolean/Phrase

+ Search results were exported from EbscoHost Platformon in a _generic bibliographic format_.  This is an unstructured text file (*.txt) that was processed using in R.

+ EbscoHost full journal listing retrieval date:  January 11, 2015 (http://www.ebscohost.com/title-lists)

# Data Pre-Processing
```{r initialization_dataprocessing, warning=FALSE, message=FALSE, eval=FALSE}
# Clear workspace
rm(list=ls())

@@

```{r initialization_analysis, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
rm(list=ls())
setwd("/Users/beperron/Git/SocialWorkResearch/Data")
source("/Users/beperron/Git/BibWrangleR/functions/ggsurv.R")
load("ebscoFullListing.R")
load("piArticles.R")
load("socialWorkJournals.R")
library(dplyr)
library(ggplot2)
library(gridExtra)
library(survival)
library(grid)
library(png)
library(reshape2)

@@

Journals excluded from search results.  _Social Work_ today excluded because it is of a magazine format.

```{r}
pi.df <- pi.df %>% filter(
        record != "Journal of Applied Social Sciences" &
        record != "Early Child Development and Care" &
        record != "The Clinical Supervisor" &
        record != "Children and Youth Services Review" &
        record != "General Hospital Psychiatry" &
        record != "Canadian Journal on Aging" &
        record != "Canadian Journal of Community Mental Health" &
        record != "Behavior Modification" &
        record != "Employee Assistance Quarterly" &
        record != "Journal of Applied Behavioral Science" & 
        record != "The Scientific Review of Mental Health Practice: Objective Investigations of Controversial and Unorthodox Claims in Clinical Psychology, Psychiatry, and Social Work" &
        record != "Systems Research and Behavioral Science" & 
        record != "Journal of Applied Rehabilitation Counseling" & 
        record != "Journal of Educational & Psychological Consultation" &
        record != "Journal of Community & Applied Social Psychology" &                                          
        record != "Evaluation and Program Planning" &                                                                  
        record != "Journal of Nonverbal Behavior" &
        record != "Brain and Cognition" &
        record != "Early Education and Development" & 
        record != "American Journal of Preventive Medicine" &
        record != "Social Work Today")
```

Two journal titles were merged because one set of records was a logical abbreviation of another set of records.

@@

pi.df$record <- ifelse(pi.df$record == "Practice", "Practice: Social Work in Action"
                , pi.df$record)
pi.df$articleID <- ifelse(pi.df$articleID == 74, 42, pi.df$articleID)

pi.df$record <- ifelse(pi.df$record == "Eugenics & Social Welfare Bull.", "Eugenics and Social Welfare Bulletin", pi.df$record)
pi.df$articleID <- ifelse(pi.df$articleID == 28, 32, pi.df$articleID)
```

# Preliminary Analyses

All analyses provided here are not in publishable format, but presented to determine the basis of a storyline.  All tables and figures will be styled according to a target journal.

## Summary of all journal titles included in the analysis

To do:

+ In the table, include the first and last years of indexing, and the numbers of articles.  
+ Write the script to output a *.csv file so that it can be included as an Appendix without any formatting.

@@

number.journals
journals.unique
socialWorkJournals <- journals.unique
#save(journals.unique, file="socialWorkJournals.R")
```

## Count of unique journal titles over time

To do:
+ Integrate the full journal listing of EbscoHost 
+ Plot overall number of journals over time 
+ Plot overall number of journals by major discipline category as defined by EbscoHost classification -- use sociology, anthropology, economics, psychology, and political science

```{r unique_titles_year, message=FALSE, comment=NA, warning=FALSE, echo=FALSE}
journals.year <- tbl_df(pi.df)

@@

journal.count <- ggplot(n.journals.year, aes(as.numeric(year), y=n, group=1)) + 
    geom_line(colour="darkblue") +
    #geom_point(colour="red") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    xlab("Publication year") + 
    ylab("Number") + 
    ggtitle("Number of Journals Over Time") + 
    scale_x_continuous(breaks=seq(1914, 2014, 5))

@@

## Growth of Social Work compared to growth of the social sciences

```{r}
ebscoFullListing.df <- ebscoFullListing.df[ebscoFullListing.df$SourceType == "Academic Journal",]
ebscoFullListing.df$IndexingStart <- format(as.Date(ebscoFullListing.df$IndexingStart, "%m/%d/%Y"), "%Y")
ebscoFullListing.df$IndexingStop <- format(as.Date(ebscoFullListing.df$IndexingStop, "%m/%d/%Y"), "%Y")
ebJournals.df <- ebscoFullListing.df %>% filter(SourceType == "Academic Journal") %>%
    select(Discipline, IndexingStart, IndexingStop, PublicationName, ISSN) %>%
    mutate(IndexingStart = as.numeric(IndexingStart), IndexingStop = as.numeric(IndexingStop)) %>%
    filter(Discipline == "Sociology" | Discipline == "Political Science & Studies"| 
           Discipline == "Psychology"| Discipline == "Economics" | Discipline == "Anthropology")

ebJournals.df.unique <- ebJournals.df %>% select(ISSN)
ebJournals.df.unique <- !duplicated(ebJournals.df.unique)
ebJournals.df <- ebJournals.df[ebJournals.df.unique, ]

ebJournals.df <- ebJournals.df[!(ebJournals.df$PublicationName %in% journals.unique),]

ebJournals.df <- select(ebJournals.df, -PublicationName, -ISSN) %>%
    mutate(journalID = c(1:nrow(ebJournals.df))) %>%
    mutate(journalID = as.factor(journalID)) %>% 
    select(-Discipline)

#ebJournals.melted <- melt(ebJournals.df)

#ebJournals.overtime.start <- ebJournals.melted %>% 
#    filter(value = !is.na(value)) %>%
#    group_by(Discipline, value) %>%
#    summarise(N.start = n())

#ebJournals.overtime.end <- ebJournals.melted %>% 
#    filter(variable == "IndexingStop"  & !is.na(value)) %>%
#    group_by(Discipline, value) %>%
#    summarise(N.end = n())

#ebjournals.joined <- left_join(ebJournals.overtime.start, ebJournals.overtime.end) %>%
#    mutate(N.end = ifelse(is.na(N.end), 0, N.end), Cumulative = cumsum(N.start)) %>%
#    mutate(N.adjusted = Cumulative - N.end)

ebJournals.melted.d <- melt(ebJournals.df)

ebJournals.overtime.start.d <- ebJournals.melted.d %>% 
    filter(value = !is.na(value)) %>%
    group_by(value) %>%
    summarise(N.start = n())

ebJournals.overtime.end.d <- ebJournals.melted.d %>% 
    filter(variable == "IndexingStop"  & !is.na(value)) %>%
    group_by(value) %>%
    summarise(N.end = n())

ebjournals.joined.d <- left_join(ebJournals.overtime.start.d, ebJournals.overtime.end.d) %>%
    mutate(N.end = ifelse(is.na(N.end), 0, N.end), Cumulative = cumsum(N.start)) %>%
    mutate(N.adjusted = Cumulative - N.end)

journal.time.reduced <- n.journals.year %>% filter(year >= 1950)
ggplot() + 
    geom_line(data = ebjournals.joined.d, aes(x=value, y=N.adjusted), group=1, colour="purple") + 
    geom_line(data = journal.time.reduced, aes(as.numeric(year), y=n), group=1, colour="darkblue")
```

@@

To do:
+ Incorporate all EbscoHost journals and the separate disciplinary journals

@@

To do:
+ Plot against major disciplinary categories

@@

To do:  
+ Fix script that performs matching -- convert all keywords to lower before matching

@@

_DO NOT INTERPRET RESULTS HERE THIS SECTION IS NOT WORKING PROPERLY_

@@

@@

select.keywords <- c("women", "leadership")
select.keywords.df <- kp.year %>% filter(tolower(keyword) %in% select.keywords)

@@

## Location of Studies

To do:
+ Determine best format / presentation of map

@@

@@

To do:
+ Determine best format / presentation of map

@@

ggplot(location.year.top, aes(x = year, y = log_n, colour = country )) + 
    geom_line() +
    ggtitle("Log Number of Articles by Countries With Highest Scholarly Ouput") +
    xlab("Publication Year") + 
    ylab("log N+1") +
    theme(legend.position=c(.210, .7))

@@

ggplot(location.year.top.no.us, aes(x = year, y = cumsum_n, colour = country )) + geom_line() +
    ggtitle("Cumulative Number of Article by Countries with Highest Scholarly Output \n(Excluding the United States)") +
    theme(legend.position=c(.210, .6))
```
## Countries represents in the social work research

To do:
+ Obtain definition of `LO` tag.  
+ Determine how many articles had LO tag (note the issue if not included from Social Science Index)

```{r fig.width=8, echo=FALSE}
Map100YearWithUS <- readPNG("/Users/beperron/Git/SocialWorkResearch/images/Map100Years.png")
grid.raster(Map100YearWithUS)

Map100YearWithoutUS <- readPNG("/Users/beperron/Git/SocialWorkResearch/images/Map100YearsWithoutUS.png")
grid.raster(Map100YearWithUS)
```

@@

```{r fig.width=8, echo=FALSE}
library(png)
library(grid)
img <- readPNG("/Users/beperron/Git/SocialWorkResearch/images/NotRepresented100Years.png")
grid.raster(img)
